---
title: "Study Diagnostics"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    number_sections: yes
    toc: yes
vignette: |
  %\VignetteIndexEntry{Study Diagnostics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

The **SelfControlledCohort** package includes a suite of diagnostics that evaluate whether the assumptions of the Self-Controlled Cohort (SCC) design hold for a given analysis. These diagnostics run automatically when `runDiagnostics = TRUE` and determine whether study results should be **unblinded** (viewed) or kept blinded until issues are resolved.

This vignette describes each diagnostic, the assumption it checks, and how results are interpreted.

# Overview

Four diagnostics are executed per target–outcome pair:

| Diagnostic Name | Assumption Tested | Default Threshold |
|---|---|---|
| **MDRR** | Adequate statistical power | MDRR ≤ 10.0 |
| **PRE_EXPOSURE** | Correct temporal ordering | Proportion ≤ 5%, p > 0.05 |
| **EVENT_DEPENDENT_OBSERVATION** | Non-informative censoring | Proportion ≤ 10% |
| **TIME_TREND_P_VALUE** | Stable baseline risk | p > 0.05 |

Default thresholds are available via `getDefaultDiagnosticThresholds()`:

```{r thresholds}
library(SelfControlledCohort)
str(getDefaultDiagnosticThresholds())
```

# Minimum Detectable Relative Risk (MDRR)

## What it checks

The MDRR quantifies the smallest rate ratio the study has 80% power to detect at α = 0.05. A high MDRR indicates that only very large effects would be detected — the study is underpowered.

## Method

The calculation uses a binary search to find the rate ratio $RR$ satisfying: $$\text{Power} = \Phi\!\left(\frac{\log(RR)}{\sqrt{1/E_1 + 1/E_0}} - z_{\alpha/2}\right) = 0.80$$ where $E_1$ and $E_0$ are the expected events in exposed and unexposed windows, using the observed baseline rate.

## Interpretation

- **MDRR ≤ 10.0** → Pass. The study has sufficient power to detect clinically relevant effects.
- **MDRR > 10.0** → Fail. The study can only detect very large effects, and estimates may be unreliable.
- **MDRR = NA** → Fail. Occurs when there are zero events or zero person-time.

## Example

```{r mdrr_examples}
# Well-powered study
computeMdrrForRateRatio(
  exposedPersonTime = 50000,
  unexposedPersonTime = 150000,
  exposedEvents = 40,
  unexposedEvents = 90
)

# Underpowered study
computeMdrrForRateRatio(
  exposedPersonTime = 500,
  unexposedPersonTime = 1500,
  exposedEvents = 3,
  unexposedEvents = 7
)
```

## Role in blinding

MDRR is the only diagnostic that affects **Tier 2 (UNBLIND)** but not **Tier 1 (UNBLIND_FOR_CALIBRATION)**. This means a low-powered study can still serve as a negative control for empirical calibration, even if its point estimate should not be viewed directly.

# Pre-Exposure Gain

## What it checks

This diagnostic detects whether outcomes occur *before* the exposure start date at a rate higher than expected. In a properly specified SCC analysis, outcomes should not systematically precede exposure.

## Why it matters

Pre-exposure outcomes suggest one or more of:

- **Confounding by indication** — the outcome (or a related condition) prompted the exposure.
- **Misspecified cohort definitions** — the exposure definition accidentally captures outcome events or vice versa.
- **Data quality issues** — incorrect temporal ordering in the source data.

Any of these violates the key SCC assumption that exposed and unexposed windows provide unbiased rate estimates.

## Method

For each target–outcome pair:

1. Count the number of persons with at least one outcome occurring before `exposure_start_date`.
2. Compute the proportion: $$p = \frac{\text{persons with pre-exposure outcome}}{\text{total persons}}$$
3. Run a one-sided binomial test (H₀: $p = 0.01$, H₁: $p > 0.01$).

## Interpretation

The diagnostic emits two rows: `PRE_EXPOSURE_PROPORTION` and `PRE_EXPOSURE_P_VALUE`.

- **Pass** if proportion ≤ 5% **and** binomial p-value > 0.05.
- **Fail** otherwise. Investigate whether the exposure and outcome definitions overlap temporally, or whether confounding by indication is present.

# Event-Dependent Observation

## What it checks

This diagnostic identifies whether the observation period ends shortly after an outcome event. If it does, the outcome may be causing censoring (e.g., the outcome leads to death or disenrollment), which biases the rate ratio.

## Why it matters

The SCC design compares rates across exposed and unexposed windows within the same person. If observation tends to end after the outcome, then:

- Outcomes near the end of observation are more likely to be observed than outcomes that would have occurred later.
- The exposed window (typically after the unexposed window) is disproportionately affected, inflating the rate ratio.

## Method

For each person with an outcome during the risk windows, the diagnostic checks whether their `observation_period_end_date` falls within 30 days after the outcome: $$\text{proportion} = \frac{\text{persons censored within 30 days of outcome}}{\text{total persons with outcome}}$$

## Interpretation

- **Proportion ≤ 10%** → Pass. Censoring after the outcome is uncommon.
- **Proportion > 10%** → Fail. A substantial fraction of patients leave observation shortly after the outcome, suggesting event-dependent censoring. Consider whether the outcome includes fatal or near-fatal events.

# Time Trend

## What it checks

This diagnostic tests whether the baseline outcome rate changes over calendar time. The SCC design implicitly assumes a stable background rate — if risks are trending upward or downward independent of exposure, the rate ratio estimate can be confounded.

## Why it matters

Calendar time trends arise from:

- **Seasonal patterns** in disease incidence (e.g., respiratory infections in winter).
- **Changes in clinical coding or diagnostic practices** over observation years.
- **Population composition changes** — the mix of patients changes over time.

If such trends exist, the rate during the unexposed window (often before exposure) and the exposed window (after) may differ for reasons unrelated to the drug.

## Method

The diagnostic aggregates outcome counts by calendar month across all persons in the analysis, then fits a Poisson GLM: $$\log\!\bigl(E[\text{count}]\bigr) = \beta_0 + \beta_1 \cdot \text{calendar\_month} + \text{offset}\!\bigl(\log(\text{person-time})\bigr)$$ The person-time offset is computed from the actual risk window durations (summing exposed and unexposed window lengths). The p-value for $\beta_1$ is tested against the threshold.

## Interpretation

- **p > 0.05** → Pass. No evidence of a calendar time trend.
- **p ≤ 0.05** → Fail. A statistically significant trend in outcome risk was detected. Consider whether seasonal adjustment or restriction to a more stable time period is appropriate.

Note: if fewer than 3 calendar months of data are available, the model cannot be fit and the diagnostic passes by default (NA).

# Tiered Blinding

The individual diagnostics feed into a two-tier blinding system that governs whether results can be viewed:

```
                     ┌──────────────────────────────────┐
                     │   All non-MDRR diagnostics pass? │
                     └──────────┬───────────────────────┘
                                │
                    ┌───── Yes ──┴── No ─────┐
                    │                        │
          ┌─────────▼──────────┐    UNBLIND_FOR_CALIBRATION = 0
          │  MDRR also passes? │          UNBLIND = 0
          └────────┬───────────┘
                   │
         ┌── Yes ──┴── No ──┐
         │                   │
    UNBLIND = 1        UNBLIND = 0
    UNBLIND_FOR_      UNBLIND_FOR_
    CALIBRATION = 1   CALIBRATION = 1
```

- **UNBLIND = 1**: All diagnostics passed. The result is suitable for direct interpretation.
- **UNBLIND_FOR_CALIBRATION = 1**: All non-power diagnostics passed. The result can be used as a negative control for empirical calibration, even if the MDRR threshold was not met.
- **Both = 0**: Core diagnostics failed. The result should remain blinded pending investigation.

# Running Diagnostics

Diagnostics are run automatically when `runDiagnostics = TRUE` (the default):

```r
runSelfControlledCohort(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = "cdm",
  exposureIds = c(1118084),
  outcomeIds = c(313217),
  databaseId = "my_db",
  resultExportPath = "results",
  runDiagnostics = TRUE
)
```

Results are saved to `scc_diagnostics_summary.csv` in the export folder.

## Customizing thresholds

```r
thresholds <- getDefaultDiagnosticThresholds()
thresholds$mdrrMaxAcceptable <- 3.0       # Allow higher MDRR
thresholds$maxPreExposureProportion <- 0.10  # Allow up to 10% pre-exposure

runSelfControlledCohort(
  ...,
  runDiagnostics = TRUE,
  diagnosticThresholds = thresholds
)
```

## Selecting specific diagnostics

```r
runSelfControlledCohort(
  ...,
  runDiagnostics = TRUE,
  diagnostics = c("mdrr", "time_trend")  # Skip pre-exposure and event-dependent
)
```

## Inspecting failures

```r
diagnostics <- read.csv("results/scc_diagnostics_summary.csv")

# Which target-outcome pairs had failures?
failures <- diagnostics[diagnostics$pass == 0 &
  !(diagnostics$diagnostic_name %in% c("UNBLIND", "UNBLIND_FOR_CALIBRATION")), ]
print(failures)
```
