---
title: "Using SelfControlledCohort"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    number_sections: yes
    toc: yes
vignette: |
  %\VignetteIndexEntry{Using SelfControlledCohort}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

The **SelfControlledCohort** package estimates the association between an exposure and an outcome by comparing the rate of outcomes during exposed time to the rate of outcomes during unexposed time within the same cohort of people. It is "self-controlled" because each person serves as their own control.

This vignette demonstrates how to run a Self-Controlled Cohort (SCC) analysis using the `Eunomia` example dataset.

# Setup and Data

First, we need to load the necessary packages and set up a connection to the `Eunomia` database. `Eunomia` provides a small, simulated Common Data Model (CDM) database for demonstration purposes.

```{r setup}
library(SelfControlledCohort)
library(Eunomia)
library(DatabaseConnector)
library(dplyr)

# Create a temporary SQLite database with Eunomia data
connectionDetails <- getEunomiaConnectionDetails()
connection <- connect(connectionDetails)

# We will use the 'main' schema for this example
cdmDatabaseSchema <- "main"
```

# Running a Basic Analysis

For a first example, let's run an analysis across **all drugs** (exposure) and **all conditions** (outcome) available in the database. 

We specify:
* `exposureTable` / `outcomeTable`: We use the standard CDM tables `drug_era` and `condition_era`.
* `exposureIds`: Left empty to include all drugs.
* `outcomeIds`: Left empty to include all conditions.

```{r basic_run}
# Define a temporary folder for results
outputFolder <- tempfile("scc_output")
dir.create(outputFolder)

# Run the analysis
runSelfControlledCohort(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = cdmDatabaseSchema,
  exposureTable = "drug_era",
  outcomeTable = "condition_era",
  resultExportPath = outputFolder,
  databaseId = 1,
  analysisId = 1,
  runDiagnostics = FALSE
)
```

Now we can read the results and inspect the estimates.

```{r read_results}
results <- read.csv(file.path(outputFolder, "scc_result.csv"))

# Display the top 5 associations by Relative Risk
results %>%
  filter(analysis_id == 1) %>%
  arrange(desc(rr)) %>%
  head(5) %>%
  select(target_cohort_id, outcome_cohort_id, rr, lb_95, ub_95, p_value) %>%
  knitr::kable(digits = 3)
```

The `rr` column shows the Incidence Rate Ratio. An IRR > 1 indicates an increased risk.

# Using Custom Cohorts

In practice, we often want to study specific cohorts defined by complex logic (e.g., "New users of NSAIDs"). We can create these cohorts and pass them to the function.

`Eunomia` includes definitions for NSAIDs and GI Bleed. Let's create them.

```{r create_cohorts}
# Create standard cohorts in the 'cohort' table:
# 1 = Celecoxib
# 2 = Diclofenac
# 3 = GI Bleed
# 4 = NSAIDs
createCohorts(connectionDetails)
```

Now run the analysis using these specific cohorts.

```{r custom_cohort_run}
runSelfControlledCohort(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = cdmDatabaseSchema,
  exposureTable = "cohort",
  outcomeTable = "cohort",
  exposureIds = c(1, 2, 4), # Celecoxib, Diclofenac, NSAIDs
  outcomeIds = c(3), # GI Bleed
  databaseId = 1,
  analysisId = 2, # Unique ID for this run
  resultExportPath = outputFolder,
  runDiagnostics = FALSE
)
```

```{r view_custom_results}
results <- read.csv(file.path(outputFolder, "scc_result.csv"))

results %>%
  filter(analysis_id == 2) %>%
  select(target_cohort_id, outcome_cohort_id, rr, p_value) %>%
  knitr::kable(digits = 3)
```

# Diagnostics and Tiered Blinding

Reliable evidence requires that the analysis assumptions are met. The package includes diagnostics to check for issues like:
*   **MDRR**: Is there enough statistical power?
*   **Time Trend**: Is the outcome risk changing over time unrelated to exposure?
*   **Pre-exposure Gain**: Are outcomes occurring unusually often just before exposure?
*   **Event-dependent Observation**: Does the outcome affect the observation period (e.g., death)?

We implement a **Tiered Blinding** system:
1.  **Tier 1 (Unblind for Calibration)**: If core diagnostics pass, the result can be used as a negative control for calibration.
2.  **Tier 2 (Unblind)**: If *all* diagnostics (including power) pass, the result is considered valid for unblinding.

Let's run the cohort analysis again with diagnostics enabled.

```{r diagnostics_run}
runSelfControlledCohort(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = cdmDatabaseSchema,
  exposureTable = "cohort",
  outcomeTable = "cohort",
  exposureIds = c(4), # NSAIDs
  outcomeIds = c(3), # GI Bleed
  analysisId = 3, # Unique ID for this run
  runDiagnostics = TRUE, # <--- Enable diagnostics
  resultExportPath = outputFolder,
  databaseId = 1
)
```

We can inspect the diagnostic summary to see the blinding status.

```{r view_blinding}
diagnostics <- read.csv(file.path(outputFolder, "scc_diagnostics_summary.csv"))

# Check blinding status for this analysis
diagnostics %>%
  filter(analysis_id == 3) %>%
  filter(diagnostic_name %in% c("UNBLIND", "UNBLIND_FOR_CALIBRATION")) %>%
  select(target_cohort_id, outcome_cohort_id, diagnostic_name, pass) %>%
  knitr::kable()
```

If `UNBLIND` is 1, the result is safe to view. If it is 0, one or more diagnostics failed. You can check which specific diagnostics failed:

```{r view_failures}
diagnostics %>%
  filter(analysis_id == 3, pass == 0) %>%
  select(diagnostic_name, diagnostic_value, pass) %>%
  knitr::kable()
```

# Empirical Calibration

Observational studies often suffer from systematic error (bias). We can use **Empirical Calibration** to adjust for this bias using negative controls—pairs of exposures and outcomes where we believe the true effect size is 1 (no effect).

In this example, we will treat **Celecoxib (1)** and **Diclofenac (2)** combined with **GI Bleed (3)** as "negative controls" purely for demonstration purposes.

We supply the `negativeControlPairs` argument. These pairs are used to fit a null distribution which is then used to calibrate the p-values and confidence intervals of our target analysis (**NSAIDs (4)** vs **GI Bleed (3)**).

```{r calibration_run}
# Define negative controls (formatted as list of vectors: c(exposureId, outcomeId))
negativeControls <- list(
  c(1, 3), # Celecoxib - GI Bleed (Demo only!)
  c(2, 3) # Diclofenac - GI Bleed (Demo only!)
)

runSelfControlledCohort(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = cdmDatabaseSchema,
  exposureTable = "cohort",
  outcomeTable = "cohort",
  exposureIds = c(4), # Target: NSAIDs
  outcomeIds = c(3), # Outcome: GI Bleed
  analysisId = 4, # Unique ID for this calibration run
  negativeControlPairs = negativeControls, # <--- Supply negative controls
  runDiagnostics = TRUE,
  resultExportPath = outputFolder,
  databaseId = 1
)
```

The results file now contains calibrated estimates (`calibrated_rr`, `calibrated_ci`, etc.).

```{r view_calibration}
results <- read.csv(file.path(outputFolder, "scc_result.csv"))

results %>%
  filter(analysis_id == 4, target_cohort_id == 4) %>%
  select(target_cohort_id, rr, calibrated_rr, calibrated_p_value) %>%
  knitr::kable(digits = 3)
```

*Note: With only 2 negative controls, the calibration will be very approximate. Real studies should use dozens of negative controls.*

# Working with the Results Database

The `SelfControlledCohort` package provides tools to create a robust results data model. This is useful for sharing results or powering a Shiny application.

In this section, we will:
1.  Create a local SQLite database to store our results.
2.  Create the standard results schema tables.
3.  Upload the CSV results we generated above into this database.
4.  Query the database to see the results.

First, let's create a new local SQLite database file for the results.

```{r results_db_creation}
resultsDbFile <- tempfile("scc_results_db", fileext = ".sqlite")
resultsConnectionDetails <- createConnectionDetails(dbms = "sqlite", server = resultsDbFile)
```

Next, we use `createResultsDataModel` to create the tables. For SQLite, we must use the 'main' schema.

```{r create_model}
createResultsDataModel(
  connectionDetails = resultsConnectionDetails,
  databaseSchema = "main"
)
```

Now we upload the results from our `outputFolder` (which contains the CSVs from all the runs above) into the database.

```{r upload_results}
# For this example, we will upload the main results and settings.
# We remove other tables to simplify the upload for this demonstration.
filesToRemove <- c("scc_diagnostics_summary.csv", "scc_stat.csv", "scc_outcome_exposure.csv")
for (f in filesToRemove) {
  if (file.exists(file.path(outputFolder, f))) {
    unlink(file.path(outputFolder, f))
  }
}

uploadResults(
  connectionDetails = resultsConnectionDetails,
  schema = "main",
  resultsFolder = outputFolder
)
```

Finally, we can connect to this results database and query it like any other OMOP CDM database.

```{r query_results_db}
resultsConn <- connect(resultsConnectionDetails)

# Query the scc_result table
resultsDb <- querySql(resultsConn, "SELECT * FROM main.scc_result WHERE analysis_id = 4")

# Show the results from the database
resultsDb %>%
  select(target_cohort_id, outcome_cohort_id, rr, calibrated_rr) %>%
  knitr::kable(digits = 3)

disconnect(resultsConn)
```

```{r cleanup, include=FALSE}
# Clean up
on.exit(disconnect(connection))
unlink(outputFolder, recursive = TRUE)
unlink(resultsDbFile)
```
