---
title: "SelfControlledCohort Results Exploration"
subtitle: "Meta-analysis results from the EvidenceSynthesis package with Reward-style benefit/risk exploration"
format: 
  html:
    page-layout: full
params:
  dbms: sqlite
  server: ""
  port: null
  user: null
  password: null
  resultsDatabaseSchema: main
  esDatabaseSchema: main
  evidenceSynthesisAnalysisId: null
  benefitThreshold: 0.5
  lowerBenefitThreshold: 0.2
  riskThreshold: 2.0
  pCutValue: 0.05
  forestPlotMaxRr: 1.1
execute:
  echo: false
  message: false
  warning: false
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE
)

library(DatabaseConnector)
library(dplyr)
library(reactable)
library(htmltools)

# Quarto renders a standalone document from the document's own directory, so a
# relative sqlite path supplied in `server` may need to be resolved against the
# document directory or the repository root before connecting.
resolveDbPath <- function(server) {
  if (is.null(server) || !nzchar(server) || file.exists(server)) {
    return(server)
  }
  candidates <- c(
    file.path(".", server),
    file.path("..", server),
    file.path("..", "..", server),
    file.path("mockResults", basename(server)),
    file.path("..", "mockResults", basename(server))
  )
  hits <- candidates[file.exists(candidates)]
  if (length(hits) > 0) hits[1] else server
}

if (params$dbms == "sqlite") {
  serverPath <- resolveDbPath(params$server)
} else {
  serverPath <- params$server
}

connectionDetails <- DatabaseConnector::createConnectionDetails(
  dbms = params$dbms,
  server = serverPath,
  port = params$port,
  user = params$user,
  password = params$password
)
connection <- DatabaseConnector::connect(connectionDetails)
```

> Note: chunks in a knitted document run in their own evaluation frames, so the connection is
> explicitly closed in the final [cleanup] chunk rather than via `on.exit`.

# Overview

This document explores SelfControlledCohort results stored in a results database. It combines two
perspectives:

1. **Per-data-source results** (`scc_result`, `scc_diagnostics_summary`, `scc_outcome_exposure`,
   `scc_analysis_setting`) with cohort names from the CohortGenerator results data model
   (`cg_cohort_definition`).
2. **Meta-analysis results** produced by the EvidenceSynthesis package in the Strategus pipeline
   (`es_analysis`, `es_scc_result`, `es_scc_diagnostics_summary`), which provide the
   meta-analysis perspective for each target/outcome/analysis combination.

## Reward-style benefit / risk exploration

Following the exploration logic of the
[Reward dashboard](https://github.com/OHDSI/Reward/blob/main/inst/sql/sql_server/dashboard/mainTable.sql),
a data source counts toward **benefit** when its calibrated relative risk is within
`[lowerBenefitThreshold, benefitThreshold]` and its calibrated p-value is below `pCutValue`
(defaults: `[0.2, 0.5]`, p < 0.05), and toward **risk** when its calibrated RR is at or above
`riskThreshold` (default: 2.0) with p < `pCutValue`. These give `benefit_count` and `risk_count`
per target/outcome/analysis combination.

## Meta-analysis perspective

The primary table is restricted to results **under meta-analysis that show a potential benefit**:

- the meta-analysis **calibrated** RR is within the benefit range and significant
  (`calibrated_rr <= benefitThreshold`, `calibrated_rr >= lowerBenefitThreshold`,
  `calibrated_p < pCutValue`), **and**
- the meta-analysis calibrated 95% CI does **not cross 1.0** (i.e. is not a null result), **and**
- the meta-analysis diagnostics pass (`es_scc_diagnostics_summary.unblind == 1`, meaning MDRR,
  EASE, I&sup2; and tau are all not FAIL).

Click on a row to inspect, for every data source, the effect estimates, event counts, benefit/risk
status, and diagnostics, together with the meta-analysis estimate and its diagnostics.

## Configuration

Render with the connection details of your results database, for example:

```bash
quarto render extras/ResultsExploration.qmd \
  -P dbms:sqlite \
  -P server:/path/to/results.sqlite \
  -P resultsDatabaseSchema:main \
  -P esDatabaseSchema:main
```

The benefit/risk thresholds can be overridden, e.g. `-P benefitThreshold:0.6 -P riskThreshold:1.5`.
The maximum RR shown on the forest-plot axes is fixed via `forestPlotMaxRr` (default 1.1).
If several Evidence Synthesis analyses exist (e.g. random- and fixed-effects), select one with
`-P evidenceSynthesisAnalysisId:<id>`; by default the first `es_analysis` with
`source_method = 'SelfControlledCohort'` is used.

### Mock data

A small synthetic mock results database is committed at
`extras/mockResults/mockResults.sqlite` (regenerate it any time with
`Rscript extras/mockResults/createMockResultsDb.R`). To render this report against it:

```bash
quarto render extras/ResultsExploration.qmd \
  -P dbms:sqlite \
  -P server:mockResults/mockResults.sqlite \
  -P resultsDatabaseSchema:main \
  -P esDatabaseSchema:main
```

Relative `server` paths are resolved against the document's directory (and the repository root),
so `-P server:extras/mockResults/mockResults.sqlite` also works.

```{r read-params, include = FALSE}
if (is.null(params$server) || nzchar(params$server) == 0) {
  stop("Please set the 'server' parameter to your results database (path for sqlite, host for postgresql).")
}
```

```{r load-tables, include = FALSE}
loadTable <- function(sql, snakeCaseToCamelCase = TRUE, ...) {
  res <- tryCatch(
    DatabaseConnector::renderTranslateQuerySql(
      connection = connection,
      sql = sql,
      snakeCaseToCamelCase = snakeCaseToCamelCase,
      ...
    ),
    error = function(e) data.frame()
  )
  as.data.frame(res)
}

sccResult <- loadTable("SELECT * FROM @schema.scc_result", schema = params$resultsDatabaseSchema)
sccDiagnostics <- loadTable("SELECT * FROM @schema.scc_diagnostics_summary", schema = params$resultsDatabaseSchema)
sccAnalysisSetting <- loadTable("SELECT * FROM @schema.scc_analysis_setting", schema = params$resultsDatabaseSchema)
sccOutcomeExposure <- loadTable("SELECT * FROM @schema.scc_outcome_exposure", schema = params$resultsDatabaseSchema)
cohortDefs <- loadTable(
  "SELECT cohort_definition_id, cohort_name FROM @cg_schema.cg_cohort_definition",
  cg_schema = params$resultsDatabaseSchema
)

esAnalyses <- loadTable("SELECT * FROM @schema.es_analysis", schema = params$esDatabaseSchema)
esSccResult <- loadTable("SELECT * FROM @schema.es_scc_result", schema = params$esDatabaseSchema)
esSccDiagnostics <- loadTable("SELECT * FROM @schema.es_scc_diagnostics_summary", schema = params$esDatabaseSchema)

cohortNames <- data.frame()
if (nrow(cohortDefs) > 0) {
  cohortNames <- cohortDefs |>
    dplyr::transmute(
      cohortId = cohortDefinitionId,
      cohortName = cohortName
    ) |>
    dplyr::distinct()
}

resolveName <- function(cohortId) {
  nm <- cohortNames$cohortName[match(cohortId, cohortNames$cohortId)]
  ifelse(is.na(nm), as.character(cohortId), nm)
}
```

```{r select-es-analysis, include = FALSE}
# Select the Evidence Synthesis analysis of interest (default: first SCC one)
if (nrow(esAnalyses) > 0) {
  esAnalyses <- esAnalyses |>
    dplyr::filter(sourceMethod == "SelfControlledCohort")
}
chosenEsId <- params$evidenceSynthesisAnalysisId
if (is.null(chosenEsId) || length(chosenEsId) == 0 ||
    !(chosenEsId %in% esAnalyses$evidenceSynthesisAnalysisId)) {
  chosenEsId <- if (nrow(esAnalyses) > 0) esAnalyses$evidenceSynthesisAnalysisId[1] else NULL
}

if (nrow(esSccResult) > 0 && !is.null(chosenEsId)) {
  esSccResult <- esSccResult |>
    dplyr::filter(evidenceSynthesisAnalysisId == chosenEsId)
}
if (nrow(esSccDiagnostics) > 0 && !is.null(chosenEsId)) {
  esSccDiagnostics <- esSccDiagnostics |>
    dplyr::filter(evidenceSynthesisAnalysisId == chosenEsId)
}

# All scc_result rows are per-database sites (meta-analysis lives in the es_ tables).
# Guard against any 'meta-analysis' marker that some pipelines write into scc_result.
siteResults <- sccResult |>
  dplyr::filter(tolower(databaseId) != "meta-analysis")
```

# Targets and outcomes of interest

Targets and outcomes are identified by cohort ids and named according to the CohortGenerator data
model (`cg_cohort_definition`). Negative controls (pairs with `true_effect_size = 1`) are flagged.

```{r prepare-targets-outcomes, include = FALSE}
targets <- data.frame()
outcomes <- data.frame()
if (nrow(sccOutcomeExposure) > 0) {
  negativeTargetIds <- unique(sccOutcomeExposure$targetCohortId[sccOutcomeExposure$trueEffectSize == 1 & !is.na(sccOutcomeExposure$trueEffectSize)])
  negativeOutcomeIds <- unique(sccOutcomeExposure$outcomeCohortId[sccOutcomeExposure$trueEffectSize == 1 & !is.na(sccOutcomeExposure$trueEffectSize)])

  targets <- sccOutcomeExposure |>
    dplyr::select(targetCohortId) |>
    dplyr::distinct() |>
    dplyr::mutate(
      targetCohortName = resolveName(targetCohortId),
      negativeControl = targetCohortId %in% negativeTargetIds
    ) |>
    dplyr::arrange(targetCohortName)

  outcomes <- sccOutcomeExposure |>
    dplyr::select(outcomeCohortId) |>
    dplyr::distinct() |>
    dplyr::mutate(
      outcomeCohortName = resolveName(outcomeCohortId),
      negativeControl = outcomeCohortId %in% negativeOutcomeIds
    ) |>
    dplyr::arrange(outcomeCohortName)
}
```

### Targets of interest (`r nrow(targets)`)

```{r targets-table}
if (nrow(targets) > 0) {
  reactable::reactable(
    targets,
    columns = list(
      targetCohortId = reactable::colDef(name = "Target cohort id"),
      targetCohortName = reactable::colDef(name = "Target cohort name"),
      negativeControl = reactable::colDef(name = "Negative control", cell = function(value) {
        if (isTRUE(value)) "Yes" else "No"
      })
    ),
    defaultPageSize = 10
  )
} else {
  cat("No exposure-outcome pairs found in `scc_outcome_exposure`.")
}
```

### Outcomes of interest (`r nrow(outcomes)`)

```{r outcomes-table}
if (nrow(outcomes) > 0) {
  reactable::reactable(
    outcomes,
    columns = list(
      outcomeCohortId = reactable::colDef(name = "Outcome cohort id"),
      outcomeCohortName = reactable::colDef(name = "Outcome cohort name"),
      negativeControl = reactable::colDef(name = "Negative control", cell = function(value) {
        if (isTRUE(value)) "Yes" else "No"
      })
    ),
    defaultPageSize = 10
  )
}
```

# Analysis settings

```{r analysis-settings}
if (nrow(sccAnalysisSetting) > 0) {
  reactable::reactable(
    sccAnalysisSetting |>
      dplyr::select(analysisId, description) |>
      dplyr::rename(`Analysis id` = analysisId, Description = description),
    defaultPageSize = 10
  )
} else {
  cat("No analysis settings found in `scc_analysis_setting`.")
}
```

# Evidence Synthesis analyses

```{r es-analyses}
if (nrow(esAnalyses) > 0) {
  esAnalysesTable <- esAnalyses |>
    dplyr::select(evidenceSynthesisAnalysisId, evidenceSynthesisDescription, sourceMethod) |>
    dplyr::mutate(selected = evidenceSynthesisAnalysisId == chosenEsId) |>
    dplyr::rename(
      `ES analysis id` = evidenceSynthesisAnalysisId,
      Description = evidenceSynthesisDescription,
      `Source method` = sourceMethod
    )
  reactable::reactable(
    esAnalysesTable,
    columns = list(
      `ES analysis id` = reactable::colDef(),
      Description = reactable::colDef(),
      `Source method` = reactable::colDef(),
      selected = reactable::colDef(name = "Selected for this report", cell = function(value) {
        if (isTRUE(value)) "Yes" else ""
      })
    ),
    defaultPageSize = 10
  )
} else {
  cat("No SelfControlledCohort Evidence Synthesis analyses found in `es_analysis`.")
}
```

# Diagnostic pass summary

Number of target/outcome pairs per analysis id that are under meta-analysis, pass the meta
diagnostics (`unblind = 1`), and show a potential benefit (within the benefit range with a CI that
does not cross 1.0).

```{r diagnostic-summary}
benefitRangeOk <- function(rr, p) {
  !is.na(rr) & !is.na(p) &
    rr <= params$benefitThreshold & rr >= params$lowerBenefitThreshold & p < params$pCutValue
}

benefitRisk <- data.frame()
if (nrow(siteResults) > 0) {
  benefitRisk <- siteResults |>
    dplyr::mutate(
      isBenefit = benefitRangeOk(calibratedRr, calibratedPValue),
      isRisk = !is.na(calibratedRr) & !is.na(calibratedPValue) &
        calibratedRr >= params$riskThreshold & calibratedPValue < params$pCutValue
    ) |>
    dplyr::group_by(analysisId, targetCohortId, outcomeCohortId) |>
    dplyr::summarise(
      benefitCount = sum(.data$isBenefit, na.rm = TRUE),
      riskCount = sum(.data$isRisk, na.rm = TRUE),
      nSources = dplyr::n_distinct(.data$databaseId),
      .groups = "drop"
    )
}

metaJoined <- data.frame()
if (nrow(esSccResult) > 0) {
  metaJoined <- esSccResult |>
    dplyr::left_join(
      esSccDiagnostics |>
        dplyr::select(targetCohortId, outcomeCohortId, analysisId,
                      mdrr, i2, tau, ease,
                      mdrrDiagnostic, i2Diagnostic, tauDiagnostic, easeDiagnostic, unblind),
      by = c("targetCohortId", "outcomeCohortId", "analysisId")
    )
}

passingMeta <- data.frame()
if (nrow(metaJoined) > 0) {
  passingMeta <- metaJoined |>
    dplyr::filter(.data$unblind == 1) |>
    dplyr::mutate(
      ciCrossesNull = .data$calibratedCi95Lb < 1 & .data$calibratedCi95Ub > 1,
      inBenefitRange = benefitRangeOk(.data$calibratedRr, .data$calibratedP)
    ) |>
    dplyr::filter(.data$inBenefitRange & !.data$ciCrossesNull)
}

if (nrow(metaJoined) > 0) {
  summaryStats <- metaJoined |>
    dplyr::group_by(analysisId) |>
    dplyr::summarise(nMetaResults = dplyr::n(), .groups = "drop") |>
    dplyr::left_join(
      metaJoined |>
        dplyr::filter(.data$unblind == 1) |>
        dplyr::group_by(analysisId) |>
        dplyr::summarise(nPassDiagnostics = dplyr::n(), .groups = "drop"),
      by = "analysisId"
    ) |>
    dplyr::left_join(
      passingMeta |>
        dplyr::group_by(analysisId) |>
        dplyr::summarise(nPotentialBenefit = dplyr::n(), .groups = "drop"),
      by = "analysisId"
    ) |>
    dplyr::mutate(
      nPassDiagnostics = dplyr::coalesce(.data$nPassDiagnostics, 0L),
      nPotentialBenefit = dplyr::coalesce(.data$nPotentialBenefit, 0L),
      pctBenefit = round(100 * .data$nPotentialBenefit / .data$nMetaResults, 1)
    ) |>
    dplyr::left_join(
      sccAnalysisSetting |>
        dplyr::select(analysisId, description),
      by = "analysisId"
    ) |>
    dplyr::arrange(dplyr::desc(.data$nPotentialBenefit), analysisId)

  reactable::reactable(
    summaryStats,
    columns = list(
      analysisId = reactable::colDef(name = "Analysis id"),
      description = reactable::colDef(name = "Description"),
      nMetaResults = reactable::colDef(name = "Meta-analysis results", align = "right"),
      nPassDiagnostics = reactable::colDef(name = "Pass meta diagnostics", align = "right"),
      nPotentialBenefit = reactable::colDef(name = "Potential benefit", align = "right"),
      pctBenefit = reactable::colDef(name = "% potential benefit", align = "right", format = reactable::colFormat(suffix = "%"))
    ),
    defaultPageSize = 10
  )
} else {
  cat("No meta-analysis results found in `es_scc_result` for the selected Evidence Synthesis analysis.")
}
```

# Results showing potential benefit

Each row is a target / outcome / analysis id combination whose meta-analysis result is within the
benefit range, is significant, does **not** cross the null (RR = 1), and passes the meta
diagnostics. `benefit_count` / `risk_count` show how many data sources fall in the benefit / risk
range, and the final column shows the meta-analysis estimate (calibrated RR from `es_scc_result`,
with I&sup2; from `es_scc_diagnostics_summary`). Forest-plot axes use a fixed maximum RR
(`forestPlotMaxRr`, default 1.1). **Click on a row** to view the grouped combination as a forest
plot with one row per data source (plus the meta-analysis), alongside a per-data-source table with
effect estimates, event counts, and diagnostics.

```{r forest-plot-helpers, include = FALSE}
fmtNum <- function(x) {
  if (length(x) == 0 || is.na(x) || !is.finite(x)) "NA" else sprintf("%.2f", x)
}
fmtCi <- function(lb, ub) {
  ok <- !is.na(lb) && is.finite(lb) && lb > 0 && !is.na(ub) && is.finite(ub) && ub > 0
  if (ok) sprintf("(%.2f - %.2f)", lb, ub) else "(NA)"
}
fmtP <- function(p) {
  if (length(p) == 0 || is.na(p)) "NA"
  else if (p < 0.001) format(p, scientific = TRUE, digits = 2)
  else sprintf("%.3f", p)
}
fmtDiagVal <- function(v) {
  if (length(v) == 0 || is.na(v)) "-"
  else if (v != 0 && abs(v) < 0.001) format(v, scientific = TRUE, digits = 2)
  else sprintf("%.3f", v)
}
hasVal <- function(v) length(v) == 1 && !is.na(v) && is.finite(v) && v > 0

# Log-RR domain for a forest plot. The maximum is fixed (default 1.1) via the
# `forestPlotMaxRr` parameter so plots share a common right edge; the minimum is
# driven by the data (the reference value RR = 1 is always included).
rowDomain <- function(rr, lb, ub, fixedMax = params$forestPlotMaxRr) {
  vals <- c(rr, lb, ub)
  vals <- vals[is.finite(vals) & vals > 0]
  vals <- c(log(vals), 0)
  lo <- min(vals)
  hi <- log(fixedMax)
  if (lo >= hi) hi <- lo + 0.5
  pad <- (hi - lo) * 0.05
  c(lo - pad, hi)
}

# Hand-rolled inline SVG forest plot (no plot file is created). Shows the estimate,
# its confidence interval whisker, and small text with the RR and CI bounds.
forestSvg <- function(rr, lb, ub, xMin, xMax, meta = FALSE, i2 = NULL,
                      piLb = NULL, piUb = NULL, width = 175, height = 52) {
  textX <- 31
  plotLeft <- 62
  plotRight <- width - 8
  yMid <- 27

  xPos <- function(v) {
    t <- (log(v) - xMin) / (xMax - xMin)
    plotLeft + t * (plotRight - plotLeft)
  }

  s <- sprintf(
    '<svg xmlns="http://www.w3.org/2000/svg" width="%d" height="%d" viewBox="0 0 %d %d">',
    width, height, width, height
  )

  # Reference line at RR = 1
  refX <- xPos(1)
  refX <- max(plotLeft, min(plotRight, refX))
  s <- paste0(s, sprintf(
    '<line x1="%.1f" y1="4" x2="%.1f" y2="%d" stroke="#cccccc" stroke-width="1" stroke-dasharray="3,2"/>',
    refX, refX, height - 4
  ))

  # Small text with RR and CI bounds
  s <- paste0(s, sprintf(
    '<text x="%.1f" y="20" font-size="9" fill="#333333" text-anchor="middle" font-family="monospace">%s</text>',
    textX, fmtNum(rr)
  ))
  s <- paste0(s, sprintf(
    '<text x="%.1f" y="33" font-size="8" fill="#777777" text-anchor="middle" font-family="monospace">%s</text>',
    textX, fmtCi(lb, ub)
  ))

  if (hasVal(rr) && hasVal(lb) && hasVal(ub)) {
    xr <- xPos(rr)
    xl <- xPos(lb)
    xu <- xPos(ub)
    # Whisker with end caps
    s <- paste0(s, sprintf(
      '<line x1="%.1f" y1="%d" x2="%.1f" y2="%d" stroke="#666666" stroke-width="2"/>',
      xl, yMid, xu, yMid
    ))
    s <- paste0(s, sprintf(
      '<line x1="%.1f" y1="%d" x2="%.1f" y2="%d" stroke="#666666" stroke-width="1.5"/>',
      xl, yMid - 5, xl, yMid + 5
    ))
    s <- paste0(s, sprintf(
      '<line x1="%.1f" y1="%d" x2="%.1f" y2="%d" stroke="#666666" stroke-width="1.5"/>',
      xu, yMid - 5, xu, yMid + 5
    ))
    if (meta) {
      # Prediction interval (wider, lighter whisker) if available
      if (hasVal(piLb) && hasVal(piUb) && hasVal(lb) && hasVal(ub)) {
        xpl <- xPos(piLb)
        xpu <- xPos(piUb)
        s <- paste0(s, sprintf(
          '<line x1="%.1f" y1="%d" x2="%.1f" y2="%d" stroke="#cba3a3" stroke-width="1.5"/>',
          xpl, yMid, xpu, yMid
        ))
      }
      s <- paste0(s, sprintf(
        '<polygon points="%.1f,%d %.1f,%d %.1f,%d %.1f,%d" fill="#B22222" stroke="#B22222"/>',
        xr, yMid - 4, xr + 4, yMid, xr, yMid + 4, xr - 4, yMid
      ))
      if (!is.null(i2) && is.finite(i2)) {
        s <- paste0(s, sprintf(
          '<text x="%.1f" y="46" font-size="8" fill="#B22222" text-anchor="middle" font-family="monospace">I&sup2; = %.0f%%</text>',
          textX, 100 * i2
        ))
      }
    } else {
      s <- paste0(s, sprintf(
        '<circle cx="%.1f" cy="%d" r="3.2" fill="#1f77b4" stroke="#ffffff" stroke-width="0.8"/>',
        xr, yMid
      ))
    }
  } else {
    s <- paste0(s, sprintf(
      '<text x="%.1f" y="15" font-size="8" fill="#999999" text-anchor="middle" font-family="monospace">not evaluated</text>',
      textX
    ))
  }

  paste0(s, '</svg>')
}

# Hand-rolled inline SVG forest plot for a grouped combination: one row per
# data source, plus a meta-analysis row (diamond). Shared fixed x-max (RR scale).
# Built as htmltools tags (not a raw HTML string) so reactable's details renderer
# displays it as an image rather than escaping it to text.
groupForestTag <- function(rr, lb, ub, label, meta = FALSE,
                           fixedMax = params$forestPlotMaxRr, width = 560) {
  dom <- rowDomain(rr, lb, ub, fixedMax = fixedMax)
  n <- length(rr)
  labelW <- 160
  textW <- 90
  plotL <- labelW + 10
  plotR <- width - textW
  rowH <- 20
  top <- 12
  height <- top + n * rowH + 6

  xPos <- function(v) {
    t <- (log(v) - dom[1]) / (dom[2] - dom[1])
    plotL + t * (plotR - plotL)
  }
  yMid <- function(i) top + (i - 1) * rowH + rowH / 2

  # Reference line at RR = 1
  refX <- xPos(1)
  refX <- max(plotL, min(plotR, refX))

  elems <- list(tags$line(
    x1 = refX, y1 = top - 2, x2 = refX, y2 = height - 2,
    stroke = "#dddddd", `stroke-width` = 1, `stroke-dasharray` = "3,2"
  ))

  for (i in seq_len(n)) {
    ym <- yMid(i)
    isMeta <- isTRUE(meta[i])
    col <- if (isMeta) "#B22222" else "#333333"

    # Label
    elems[[length(elems) + 1]] <- tags$text(
      x = labelW - 6, y = ym + 4, `font-size` = 11, fill = col,
      `text-anchor` = "end", `font-family` = "sans-serif", label[i]
    )
    # RR and CI text on the right
    elems[[length(elems) + 1]] <- tags$text(
      x = plotR + 8, y = ym + 4, `font-size` = 10, fill = col,
      `font-family` = "monospace",
      sprintf("%s %s", fmtNum(rr[i]), fmtCi(lb[i], ub[i]))
    )

    if (hasVal(rr[i]) && hasVal(lb[i]) && hasVal(ub[i])) {
      xr <- xPos(rr[i])
      xl <- xPos(lb[i])
      xu <- xPos(ub[i])
      elems[[length(elems) + 1]] <- tags$line(
        x1 = xl, y1 = ym, x2 = xu, y2 = ym, stroke = "#666666", `stroke-width` = 1.5
      )
      elems[[length(elems) + 1]] <- tags$line(
        x1 = xl, y1 = ym - 3.5, x2 = xl, y2 = ym + 3.5, stroke = "#666666", `stroke-width` = 1
      )
      elems[[length(elems) + 1]] <- tags$line(
        x1 = xu, y1 = ym - 3.5, x2 = xu, y2 = ym + 3.5, stroke = "#666666", `stroke-width` = 1
      )
      if (isMeta) {
        elems[[length(elems) + 1]] <- tags$polygon(
          points = sprintf(
            "%.1f,%.1f %.1f,%.1f %.1f,%.1f %.1f,%.1f",
            xr, ym - 3.5, xr + 3.5, ym, xr, ym + 3.5, xr - 3.5, ym
          ),
          fill = "#B22222", stroke = "#B22222"
        )
      } else {
        elems[[length(elems) + 1]] <- tags$circle(
          cx = xr, cy = ym, r = 3, fill = "#1f77b4", stroke = "#ffffff", `stroke-width` = 0.6
        )
      }
    }
  }

  tags$svg(
    xmlns = "http://www.w3.org/2000/svg",
    width = width,
    height = height,
    viewBox = sprintf("0 0 %d %d", width, height),
    elems
  )
}
```

```{r build-main-table, include = FALSE}
siteIds <- sort(unique(siteResults$databaseId))

if (nrow(passingMeta) > 0 && length(siteIds) > 0) {
  passingMeta <- passingMeta |>
    dplyr::left_join(benefitRisk, by = c("analysisId", "targetCohortId", "outcomeCohortId"))

  siteWide <- siteResults |>
    dplyr::inner_join(
      passingMeta |>
        dplyr::select(analysisId, targetCohortId, outcomeCohortId) |>
        dplyr::distinct(),
      by = c("analysisId", "targetCohortId", "outcomeCohortId")
    )

  reportRows <- passingMeta |>
    dplyr::select(analysisId, targetCohortId, outcomeCohortId,
                  benefitCount, riskCount, nSources,
                  calibratedRr, calibratedCi95Lb, calibratedCi95Ub, calibratedP,
                  pi95Lb, pi95Ub, nDatabases,
                  numOutcomesExposed, numOutcomesUnexposed, numExposures,
                  mdrr, i2, tau, ease,
                  mdrrDiagnostic, i2Diagnostic, tauDiagnostic, easeDiagnostic, unblind)
  reportRows$sites <- lapply(seq_len(nrow(reportRows)), function(i) {
    siteWide[siteWide$analysisId == reportRows$analysisId[i] &
               siteWide$targetCohortId == reportRows$targetCohortId[i] &
               siteWide$outcomeCohortId == reportRows$outcomeCohortId[i], ]
  })

  # Compact HTML for a set of diagnostic rows
  diagHtmlString <- function(diags) {
    if (nrow(diags) == 0) return("")
    lines <- vapply(seq_len(nrow(diags)), function(j) {
      d <- diags[j, ]
      status <- if (is.na(d$pass)) "NOT EVALUATED" else if (d$pass == 1) "PASS" else "FAIL"
      color <- switch(status, PASS = "#27ae60", FAIL = "#c0392b", "#7f8c8d")
      sprintf("%s: %s <span style='color:%s;font-weight:600;'>%s</span>",
              d$diagnosticName, fmtDiagVal(d$diagnosticValue), color, status)
    }, character(1))
    paste(lines, collapse = "<br/>")
  }

  metaDiagStatus <- function(x) {
    if (length(x) == 0 || is.na(x) || nchar(x) == 0) "NOT EVALUATED" else x
  }
  metaDiagColor <- function(x) {
    switch(metaDiagStatus(x), PASS = "#27ae60", FAIL = "#c0392b", "#7f8c8d")
  }
  metaDiagHtml <- function(r) {
    rows <- list(
      sprintf("MDRR: %s <span style='color:%s;font-weight:600;'>%s</span>",
              fmtDiagVal(r$mdrr), metaDiagColor(r$mdrrDiagnostic), metaDiagStatus(r$mdrrDiagnostic)),
      sprintf("EASE: %s <span style='color:%s;font-weight:600;'>%s</span>",
              fmtDiagVal(r$ease), metaDiagColor(r$easeDiagnostic), metaDiagStatus(r$easeDiagnostic)),
      sprintf("I&sup2;: %s <span style='color:%s;font-weight:600;'>%s</span>",
              fmtDiagVal(r$i2), metaDiagColor(r$i2Diagnostic), metaDiagStatus(r$i2Diagnostic)),
      sprintf("tau: %s <span style='color:%s;font-weight:600;'>%s</span>",
              fmtDiagVal(r$tau), metaDiagColor(r$tauDiagnostic), metaDiagStatus(r$tauDiagnostic))
    )
    paste(rows, collapse = "<br/>")
  }

  getSiteDiags <- function(e) {
    diags <- data.frame()
    if (nrow(sccDiagnostics) > 0) {
      diags <- sccDiagnostics |>
        dplyr::filter(
          analysisId == e$analysisId,
          targetCohortId == e$targetCohortId,
          outcomeCohortId == e$outcomeCohortId,
          databaseId == e$databaseId,
          !(diagnosticName %in% c("UNBLIND", "UNBLIND_FOR_CALIBRATION", "EASE"))
        )
      ease <- sccDiagnostics |>
        dplyr::filter(
          analysisId == e$analysisId,
          targetCohortId == e$targetCohortId,
          outcomeCohortId == 0,
          databaseId == e$databaseId,
          diagnosticName == "EASE"
        )
      diags <- dplyr::bind_rows(diags, ease)
    }
    diags
  }

  # Per-database rows shown when a row is expanded
  buildDbTable <- function(est, r) {
    siteRowsTbl <- lapply(seq_len(nrow(est)), function(i) {
      e <- est[i, ]
      diags <- getSiteDiags(e)
      data.frame(
        database = e$databaseId,
        rr = fmtNum(e$calibratedRr),
        ci = fmtCi(e$calibratedLb95, e$calibratedUb95),
        p = fmtP(e$calibratedPValue),
        exposed = as.character(e$numOutcomesExposed),
        unexposed = as.character(e$numOutcomesUnexposed),
        exposures = as.character(e$numExposures),
        benefit = if (benefitRangeOk(e$calibratedRr, e$calibratedPValue)) "Yes" else "No",
        risk = if (!is.na(e$calibratedRr) && !is.na(e$calibratedPValue) &&
                     e$calibratedRr >= params$riskThreshold && e$calibratedPValue < params$pCutValue) "Yes" else "No",
        diagnostics = diagHtmlString(diags),
        stringsAsFactors = FALSE
      )
    })
    dbTbl <- dplyr::bind_rows(siteRowsTbl)
    metaRow <- data.frame(
      database = "Meta-analysis",
      rr = fmtNum(r$calibratedRr),
      ci = fmtCi(r$calibratedCi95Lb, r$calibratedCi95Ub),
      p = fmtP(r$calibratedP),
      exposed = as.character(r$numOutcomesExposed),
      unexposed = as.character(r$numOutcomesUnexposed),
      exposures = as.character(r$numExposures),
      benefit = if (benefitRangeOk(r$calibratedRr, r$calibratedP)) "Yes" else "No",
      risk = "No",
      diagnostics = metaDiagHtml(r),
      stringsAsFactors = FALSE
    )
    rbind(dbTbl, metaRow)
  }

  # Expandable details: grouped forest plot + per-database table
  buildDetails <- function(est, r) {
    groupPlot <- groupForestTag(
      rr = c(est$calibratedRr, r$calibratedRr),
      lb = c(est$calibratedLb95, r$calibratedCi95Lb),
      ub = c(est$calibratedUb95, r$calibratedCi95Ub),
      label = c(est$databaseId, "Meta-analysis"),
      meta = c(rep(FALSE, nrow(est)), TRUE)
    )
    dbTbl <- buildDbTable(est, r)
    tblWidget <- reactable::reactable(
      dbTbl,
      columns = list(
        database = reactable::colDef(name = "Data source", minWidth = 140),
        rr = reactable::colDef(name = "Calibrated RR"),
        ci = reactable::colDef(name = "95% CI"),
        p = reactable::colDef(name = "p-value"),
        exposed = reactable::colDef(name = "Outcomes exposed"),
        unexposed = reactable::colDef(name = "Outcomes unexposed"),
        exposures = reactable::colDef(name = "Exposures"),
        benefit = reactable::colDef(name = "Benefit"),
        risk = reactable::colDef(name = "Risk"),
        diagnostics = reactable::colDef(name = "Diagnostics", html = TRUE, minWidth = 300)
      ),
      defaultPageSize = nrow(dbTbl),
      highlight = TRUE,
      bordered = TRUE
    )
    htmltools::tagList(
      groupPlot,
      htmltools::tags$div(style = "margin-top:8px;", tblWidget)
    )
  }

  reportDfList <- lapply(seq_len(nrow(reportRows)), function(i) {
    r <- reportRows[i, ]
    est <- r$sites[[1]]
    dom <- rowDomain(r$calibratedRr, r$calibratedCi95Lb, r$calibratedCi95Ub)
    out <- list(
      name = sprintf(
        "%s / %s / analysis %d",
        resolveName(r$targetCohortId),
        resolveName(r$outcomeCohortId),
        r$analysisId
      ),
      benefitCount = r$benefitCount,
      riskCount = r$riskCount,
      nSources = r$nSources
    )
    out$meta <- forestSvg(r$calibratedRr, r$calibratedCi95Lb, r$calibratedCi95Ub,
                          dom[1], dom[2], meta = TRUE, i2 = r$i2,
                          piLb = r$pi95Lb, piUb = r$pi95Ub)
    out$details <- buildDetails(est, r)
    out
  })

  reportDf <- do.call(rbind, lapply(reportDfList, function(x) {
    as.data.frame(x[setdiff(names(x), "details")], stringsAsFactors = FALSE, check.names = FALSE)
  }))
  detailsList <- lapply(reportDfList, function(x) x$details)
}
```

```{r render-results}
if (nrow(passingMeta) > 0 && length(siteIds) > 0) {
  columns <- list(
    name = reactable::colDef(name = "Target / Outcome / Analysis", minWidth = 260),
    benefitCount = reactable::colDef(name = "Sources in benefit range", align = "right"),
    riskCount = reactable::colDef(name = "Sources in risk range", align = "right"),
    nSources = reactable::colDef(name = "Data sources", align = "right"),
    meta = reactable::colDef(name = "Meta-analysis (EvidenceSynthesis)", html = TRUE, width = 180, align = "center")
  )

  reactable::reactable(
    reportDf,
    columns = columns,
    details = function(index) {
      detailsList[[index]]
    },
    defaultPageSize = 15,
    filterable = TRUE,
    searchable = TRUE,
    defaultSorted = list(benefitCount = "desc", name = "asc"),
    wrap = FALSE,
    highlight = TRUE,
    bordered = TRUE,
    striped = TRUE
  )
} else {
  cat("No meta-analysis results show a potential benefit within the configured range.")
}
```

```{r cleanup, include = FALSE}
if (exists("connection") && DatabaseConnector::dbIsValid(connection)) {
  DatabaseConnector::disconnect(connection)
}
```
